---
title: "February 2022 Waste Audit"
author: ""
subtitle: Duke Sustainability Task Force
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(infer)
library(car)
library(stringr)
library(robotstxt)
library(rvest)
library(httr)
library(tinytex)
library(sf)
```

```{r upload-and-clean-data}
waste_audit <- Spring_2022_February_Waste_Audit_Data_Sheet_xlsx_Figures_1_ %>% 
  select(-c(notes_contaminants, notes_stream)) %>% 
  slice(1:41, 45, 47, 48) %>% 
  select(-prop_pieces_trash, -prop_pieces_recycle, -prop_pieces_compost, -prop_trash_weight, -prop_recycle_weight, -prop_compost_weight) %>% 
  mutate(prop_pieces_trash = pieces_trash / total_piece_volume,
         prop_pieces_recycle = pieces_recycle / total_piece_volume,
         prop_pieces_compost = pieces_compost / total_piece_volume,
         prop_trash_weight = weight_trash_lbs / total_can_weight,
         prop_recycle_weight = weight_recycle_lbs / total_can_weight,
         prop_compost_weight = weight_compost_lbs / total_can_weight)

view(waste_audit)
```
# Methodology

## Content Comparisons
- average weight of a piece of trash, composting, and recycling item
```{r average-piece-weight}
waste_audit %>% 
  summarize(total_trash_pieces = sum(pieces_trash),
            total_recycle_pieces = sum(pieces_recycle),
            total_compost_pieces = sum(pieces_compost),
            total_trash_weight = sum(weight_trash_lbs), 
            total_recycle_weight = sum(weight_recycle_lbs),
            total_compost_weight = sum(weight_compost_lbs)) %>% 
  summarize(average_trash_weight = total_trash_weight / total_trash_pieces,
            average_recycle_weight = total_recycle_weight / total_recycle_pieces,
            average_compost_weight = total_compost_weight / total_compost_pieces)
```

- comparative starting volume of pieces/weight between trash, recycle, and
compost streams
```{r starting-volumes}
waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(avg_starting_weight = mean(total_can_weight),
            avg_starting_pieces = mean(total_piece_volume))
```

SOMETHING WRONG WITH THIS BAR GRAPH - TRASH & RECYCLING SHOULD BE THE TALLEST
```{r bar-plot-original-stream-piece-comparison}
data <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(total_trash = sum(pieces_trash),
            total_recycle = sum(pieces_recycle),
            total_compost = sum(pieces_compost)) %>% 
  select(total_trash, total_recycle, total_compost)
data <- as.matrix(data)
colnames(data) <- c("trash", "recycle", "compost")
rownames(data) <- c("Recycle", "Trash", "Compost")
color <- c("black", "#3265ab", "#2ca816")
barplot(t(data), beside = FALSE, col = color, ylim = c(0, 2500),
        main = "Actual Waste Composition of Trash, Recycling, and Compost (in pieces)", xlab = "Original Stream Designation", ylab = "Total Pieces")
  legend("topleft", colnames(data), pch = 15, col = color, horiz = T, bty = "n")
# source: https://stackoverflow.com/questions/61506408/grouped-barchart-in-r-with-4-variables
```

## Quantifying Missorted Waste
```{r missorted}
waste_audit <- waste_audit %>% 
  mutate(perc_missorted_c = case_when(original_stream == "Trash" ~ (pieces_recycle + pieces_compost) / total_piece_volume, original_stream == "Recycling" ~ (pieces_trash + pieces_compost) / total_piece_volume, original_stream == "Compost" ~ (pieces_trash + pieces_recycle) / total_piece_volume))
```

BELOW IS UNDONE
```{r bar-plot-original-stream-percent}
makeup <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(average_trash = mean(prop_pieces_trash),
            average_recycle = mean(prop_pieces_recycle),
            average_compost = mean(prop_pieces_compost)) %>% 
  select(average_trash, average_recycle, average_compost)
makeup <- as.matrix(makeup)
colnames(makeup) <- c("trash", "recycle", "compost")
rownames(makeup) <- c("Recycle", "Trash", "Compost")
color <- c("black", "#3265ab", "#2ca816")
barplot(t(makeup), beside = FALSE, col = color, ylim = c(0, 1.1),
        main = "Percent Actual Waste Composition in Trash, Recycling, and Compost", xlab = "Original Stream Designation", ylab = "Percent")
legend("topleft", colnames(makeup), pch = 15, col = color, horiz = T, bty = "n")
```

- percent of trash that is missorted

```{r missorted-trash}
pie_trash <- waste_audit %>% 
  mutate(missorted = pieces_recycle + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_trash),
            missorted = sum(missorted)) %>% 
  slice(2)
pie_trash
slices <- c(183, 261)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Trash")

# source: https://www.statmethods.net/graphs/pie.html
```

- percent of recycling that is missorted
    
```{r missorted-recycle}
pie_recycle <- waste_audit %>% 
  mutate(missorted = pieces_trash + pieces_compost) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_recycle),
            missorted = sum(missorted)) %>% 
  slice(1)
pie_recycle
slices <- c(31, 434)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Recycle")
```

- percent of compost that is missorted
    
```{r missorted-recycle}
pie_compost <- waste_audit %>% 
  mutate(missorted = pieces_trash + pieces_recycle) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(pieces_compost),
            missorted = sum(missorted)) %>% 
  slice(1)
pie_compost
slices <- c(282, 183)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Compost")
```

## Campus Compost
- percent of trash and recycling that could be compost
    HAVE NOT LOOKED AT THE ONES BELOW
```{r compost-in-trash}
compostable_trash <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable)) %>% 
  slice(2)
compostable_trash
slices <- c(1274, 920)
lbls <- c("compostable", "noncompostable")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Trash that is Actually Compostable")
```

```{r compost-in-recycling}
compostable_recycle <- waste_audit %>% 
  mutate(noncompostable = pieces_recycle + pieces_trash) %>% 
  group_by(original_stream) %>% 
  summarize(compostable = sum(pieces_compost),
            noncompostable = sum(noncompostable)) %>% 
  slice(1)
compostable_recycle
slices <- c(320, 343)
lbls <- c("compostable", "noncompostable")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "black"), main = "Percent of Recycling that is Actually Compostable")
```

- volume of compostables
    bar graph, value
    
```{r compost-volume-values}
volume_bar <- waste_audit %>% 
  summarize(total_pieces_compost = sum(pieces_compost),
            total_weight_compost = sum(weight_compost),
            total_pieces_audited = sum(total_piece_volume),
            total_weight_audited = sum(total_can_weight),
            perc_pieces_compost = total_pieces_compost / total_pieces_audited * 100,
            perc_weight_compost = total_weight_compost / total_weight_audited * 100)
makeup <- as.matrix(volume_bar)
volume_bar <- volume_bar %>% 
  select(perc_pieces_compost, perc_weight_compost)
colnames(volume_bar) <- c("Pieces", "Weight (lbs)")
color <- c("#2ca816", "#8cd9b3")
barplot(t(volume_bar), beside = TRUE, col = color, ylim = c(0, 100),
        main = "Percent of Total Waste that is Actually Compostable by Piece Count and Weight", xlab = "Method of Quantifying Waste", ylab = "Percent")
legend("topleft", colnames(volume_bar), pch = 15, col = color, horiz = T, bty = "n")
```

## Differences in Location
CREATE CODE FOR THIS SECTION

separate based on Loop, Zweli's, and BC Plaza
summary amounts -- how much waste sorted
contamination rates

# Bootstrap Sample

null hypothesis and alternative hypothesis

bootstrap sample

```{r bootstrap-sample-with-compost}
set.seed(405)
boot_missorted <- waste_audit %>% 
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-with-compost}
ci_boot_missorted <- get_ci(boot_missorted, level = 0.95)
ci_boot_missorted
```

bootstrap distribution

```{r bootstrap-vis-with-compost}
visualize(boot_missorted) +
  shade_ci(ci_boot_missorted) +
  labs(title = "Simulated Bootstrap Distribution", x = "Sample Mean", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-with-compost, eval = FALSE}
mean(waste_audit$perc_missorted_c) + c(-1, 1) * 2.05 * sd(waste_audit$perc_missorted_c) / sqrt(22)
```

tdist: 0.7146358, 0.7902705

# Discussion

- error: human error in sorting
- error: some contamination spreads (e.g. containers filled with liquid)
- discuss time of day and year that the data was collected
- note: this does not account for all waste; waste disposed of in dorms, WU conveyor belt, BC, library, quad trash bins, gym (Red Mango), East Campus (Marketplace, dorms)
- semi-random sampling (all recycling bins chosen, which dictated our choices in picking from the trash bins
- error: subjectivity in what it means to be proximal to an exit door, stairwell mouth, etc.