---
title: "February 2022, Weight"
author: ''
subtitle: Duke Sustainability Task Force
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---

```{r setup, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, 
                      message = FALSE, warning = FALSE,
                      fig.height = 6, fig.width = 9, fig.align = "center")
```

```{r packages}
library(tidyverse)
library(infer)
library(car)
library(stringr)
library(robotstxt)
library(rvest)
library(httr)
library(tinytex)
```

```{r upload-and-clean-data}
waste_audit <- read.csv(file = 'feb2022data.csv') %>% 
  select(-c(notes_contaminants, notes_stream)) %>% 
  slice(1:41, 45, 47, 48) %>% 
  select(-prop_pieces_trash, -prop_pieces_recycle, -prop_pieces_compost, -prop_trash_weight, -prop_recycle_weight, -prop_compost_weight) %>% 
  mutate(prop_pieces_trash = pieces_trash / total_piece_volume,
         prop_pieces_recycle = pieces_recycle / total_piece_volume,
         prop_pieces_compost = pieces_compost / total_piece_volume,
         prop_trash_weight = weight_trash_lbs / total_can_weight,
         prop_recycle_weight = weight_recycle_lbs / total_can_weight,
         prop_compost_weight = weight_compost_lbs / total_can_weight) %>%
  mutate(restaurant = case_when(
    can_number %in% c(23, 24, 25, 26, 27, 28) ~ "Zwelis",
    can_number %in% c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22) ~ "Loop",
    can_number %in% c(38, 39, 40, 30, 39, 31, 29, 47, 48, 45, 41, 35, 36, 37, 34, 32, 33) ~ "Plaza"
  ))
```

## Content Comparisons
- average weight of a piece of trash, composting, and recycling item
```{r average-piece-weight}
waste_audit %>% 
  summarize(total_trash_pieces = sum(pieces_trash),
            total_recycle_pieces = sum(pieces_recycle),
            total_compost_pieces = sum(pieces_compost),
            total_trash_weight = sum(weight_trash_lbs), 
            total_recycle_weight = sum(weight_recycle_lbs),
            total_compost_weight = sum(weight_compost_lbs),
            average_trash_weight = total_trash_weight / total_trash_pieces,
            average_recycle_weight = total_recycle_weight / total_recycle_pieces,
            average_compost_weight = total_compost_weight / total_compost_pieces)
```

- comparative starting volume of pieces/weight between trash, recycle, and
compost streams
```{r starting-volumes}
waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(avg_starting_weight = mean(total_can_weight),
            avg_starting_pieces = mean(total_piece_volume))
```

```{r bar-plot-original-stream-weight-comparison}
data <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(total_trash = sum(weight_trash_lbs),
            total_recycle = sum(weight_recycle_lbs),
            total_compost = sum(weight_compost_lbs)) %>% 
  select(total_trash, total_recycle, total_compost)
data <- as.matrix(data)
colnames(data) <- c("trash", "recycle", "compost")
rownames(data) <- c("Compost", "Recycle", "Trash")
color <- c("black", "#3265ab", "#2ca816")
barplot(t(data), beside = FALSE, col = color, ylim = c(0, 140),
        main = "Actual Waste Composition of Trash, Recycling, and Compost (by weight)", xlab = "Original Stream Designation", ylab = "Total Weight (lbs)")
  legend("topleft", colnames(data), pch = 15, col = color, horiz = T, bty = "n")
# source: https://stackoverflow.com/questions/61506408/grouped-barchart-in-r-with-4-variables
```

## Quantifying Missorted Waste
```{r missorted}
waste_audit <- waste_audit %>% 
  mutate(perc_missorted_c = case_when(original_stream == "trash" ~ (weight_recycle_lbs + weight_compost_lbs) / total_can_weight, original_stream == "recycling" ~ (weight_trash_lbs + weight_compost_lbs) / total_can_weight, original_stream == "compost" ~ (weight_trash_lbs + weight_recycle_lbs) / total_can_weight),
         missorted_count = case_when(original_stream == "trash" ~ (weight_recycle_lbs + weight_compost_lbs), original_stream == "recycling" ~ (weight_trash_lbs + weight_compost_lbs), original_stream == "compost" ~ (weight_trash_lbs + weight_recycle_lbs)),
         sorted_count = total_can_weight - missorted_count)
```

```{r bar-plot-original-stream-percent}
makeup <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(average_trash = mean(prop_trash_weight),
            average_recycle = mean(prop_recycle_weight),
            average_compost = mean(prop_compost_weight)) %>% 
  select(average_trash, average_recycle, average_compost)
makeup <- as.matrix(makeup)
colnames(makeup) <- c("trash", "recycle", "compost")
rownames(makeup) <- c("Compost", "Recycle", "Trash")
color <- c("black", "#3265ab", "#2ca816")
barplot(t(makeup), beside = FALSE, col = color, ylim = c(0, 1.1),
        main = "Percent Actual Waste Composition in Trash, Recycling, and Compost", xlab = "Original Stream Designation", ylab = "Proportion")
legend("topleft", colnames(makeup), pch = 15, col = color, horiz = T, bty = "n")
```

- percent of total waste that is missorted

```{r missorted-total-waste}
pie_total <- waste_audit %>% 
  summarize(sorted = sum(sorted_count),
            missorted = sum(missorted_count))
pie_total
slices <- c(50.5, 138.7)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Overall Contamination for All Streams")
# source: https://www.statmethods.net/graphs/pie.html
```

- Bootstrap Sample OVERALL WASTE

bootstrap sample

```{r bootstrap-sample-with-compost}
set.seed(201)
boot_missorted <- waste_audit %>% 
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-with-compost}
ci_boot_missorted <- get_ci(boot_missorted, level = 0.95)
ci_boot_missorted
```

bootstrap distribution

```{r bootstrap-vis-with-compost}
visualize(boot_missorted) +
  shade_ci(ci_boot_missorted) +
  labs(title = "Simulated Bootstrap Distribution of Overall Missorted Waste", x = "Sample Mean Missorted Overall Waste", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-with-compost, eval = FALSE}
mean(waste_audit$perc_missorted_c) + c(-1, 1) * 2.05 * sd(waste_audit$perc_missorted_c) / sqrt(44)
```

tdist: 0.5465229, 0.7112063

-percent of trash and recycling that is missorted (so we can compare last year)

```{r missorted-trash-plus-recycling}
pie_tandr <- waste_audit %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(sorted_count),
            missorted = sum(missorted_count)) %>%
  slice(2:3)
pie_tandr <- pie_tandr %>%
  summarize(sorted = sum(sorted),
            missorted = sum(missorted))
pie_tandr
slices <- c(36, 130)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Overall Contamination in both Trash and Recycling")
```

- Bootstrap Sample TRASH + RECYCLING ONLY

bootstrap sample

```{r bootstrap-sample-trash-plus-recycling}
boot_table_tr <- waste_audit %>%
  slice(1:8, 11:19, 22:24, 27:30, 32, 34:36, 38:41)
set.seed(201)
boot_missorted_tr <- boot_table_tr %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-trash-plus-recycling}
ci_boot_missorted_tr <- get_ci(boot_missorted_tr, level = 0.95)
ci_boot_missorted_tr
```

bootstrap distribution

```{r bootstrap-vis-trash-plus-recycling}
visualize(boot_missorted_tr) +
  shade_ci(ci_boot_missorted_tr) +
  labs(title = "Simulated Bootstrap Distribution of Overall Trash & Recycling", x = "Sample Mean Missorted Trash & Recycling", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-trash-plus-recycling, eval = FALSE}
mean(boot_table_tr$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_tr$perc_missorted_c) / sqrt(32)
```

tdist: 0.6698542, 0.7857089


- percent of trash that is missorted

```{r missorted-trash}
pie_trash <- waste_audit %>% 
  mutate(missorted = weight_recycle_lbs + weight_compost_lbs) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(weight_trash_lbs),
            missorted = sum(missorted)) %>%
  slice(3)
pie_trash
slices <- c(30.8, 110.8)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Trash")
```

- Bootstrap Sample TRASH ONLY

bootstrap sample

```{r bootstrap-sample-trash}
boot_table_trash <- waste_audit %>%
  slice(1:6, 11:17, 22:24, 27:29, 32, 35, 38:41)
set.seed(201)
boot_missorted_trash <- boot_table_trash %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-trash}
ci_boot_missorted_trash <- get_ci(boot_missorted_trash, level = 0.95)
ci_boot_missorted_trash
```

bootstrap distribution

```{r bootstrap-vis-trash}
visualize(boot_missorted_trash) +
  shade_ci(ci_boot_missorted_trash) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Trash", x = "Sample Mean Missorted Trash", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-trash, eval = FALSE}
mean(boot_table_trash$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_trash$perc_missorted_c) / sqrt(25)
```

tdist: 0.6506066, 0.7820920

- percent of recycling that is missorted
    
```{r missorted-recycle}
pie_recycle <- waste_audit %>% 
  mutate(missorted = weight_trash_lbs + weight_compost_lbs) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(weight_recycle_lbs),
            missorted = sum(missorted)) %>%
  slice(2)
pie_recycle
slices <- c(5.2, 19.2)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Recycle")
```

- Bootstrap Sample RECYCLING ONLY

bootstrap sample

```{r bootstrap-sample-recycling}
boot_table_recycling <- waste_audit %>%
  slice(7:8, 18:19, 30, 34, 36)
set.seed(201)
boot_missorted_recycling <- boot_table_recycling %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-recycling}
ci_boot_missorted_recycling <- get_ci(boot_missorted_recycling, level = 0.95)
ci_boot_missorted_recycling
```

bootstrap distribution

```{r bootstrap-vis-recycling}
visualize(boot_missorted_recycling) +
  shade_ci(ci_boot_missorted_recycling) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Recycling", x = "Sample Mean Missorted Recycling", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-recycling, eval = FALSE}
mean(boot_table_recycling$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_recycling$perc_missorted_c) / sqrt(7)
```

tdist:0.6420546, 0.8951677

- percent of compost that is missorted
    
```{r missorted-compost}
pie_compost <- waste_audit %>% 
  mutate(missorted = weight_trash_lbs + weight_recycle_lbs) %>% 
  group_by(original_stream) %>% 
  summarize(sorted = sum(weight_compost_lbs),
            missorted = sum(missorted)) %>% 
  slice(1)
pie_compost
slices <- c(14.5, 8.7)
lbls <- c("properly sorted", "missorted")
percent <- slices / sum(slices) * 100
percent
pct <- round(percent)
lbls <- paste(lbls, pct)
lbls <-paste(lbls, "%", sep = " ")
pie(slices, labels = lbls, col = c("#2ca816", "#cc2900"), main = "Missorted Compost")
```

- Bootstrap Sample COMPOST ONLY

bootstrap sample

```{r bootstrap-sample-compost}
boot_table_compost <- waste_audit %>%
  slice(9, 10, 20, 21, 25, 26, 31, 33, 37, 42, 43, 44)
set.seed(201)
boot_missorted_compost <- boot_table_compost %>%
  specify(response = perc_missorted_c) %>% 
  generate(reps = 10000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

confidence interval

```{r confidence-interval-compost}
ci_boot_missorted_compost <- get_ci(boot_missorted_compost, level = 0.95)
ci_boot_missorted_compost
```

bootstrap distribution

```{r bootstrap-vis-compost}
visualize(boot_missorted_compost) +
  shade_ci(ci_boot_missorted_compost) +
  labs(title = "Simulated Bootstrap Distribution of Missorted Compost", x = "Sample Mean Missorted Compost", y = "Count")
```

$\bar{x} \pm t^*\frac{s}{\sqrt{n}}$

more conservative than normal distribution; weight on tail ends

```{r t-distribution-compost, eval = FALSE}
mean(boot_table_compost$perc_missorted_c) + c(-1, 1) * 2.05 * sd(boot_table_compost$perc_missorted_c) / sqrt(12)
```

tdist: 0.1769003, 0.5532716
